---
title: "R Notebook"
output: html_document
---

```{r}
library(psych)
library(mgcv)
library(scales)
library(latex2exp)
library(knnflex)
library(DiagrammeR)
library(tidyverse)
library(utils)
library(Metrics)
library(dcurves)
library(glmnet)
library(xgboost)
library(mgcv)
library(caret)
library(ROCR)
library(plyr)
library(partykit)
library(randomForest)
library(pROC)
library(e1071)
library(class)
library(car)
library(rpart)
library(rpart.plot)
library(cutpointr)
library(EnvStats)

# Set the working directory
setwd("C:/Ngoc/MastersThesis/Draft")


# Load the .csv file
df <- read.csv('diabetes.csv')
head(df)
```

#### Preprocessing data:

```{r}
str(df)
# => The dataset has 768 observations and 9 variables. All variables are in numerical format
```

```{r}
# Since Outcome is supposed to be a categorical variable, we need to convert it into a factor 
df$Outcome <- as.factor(df$Outcome) 

str(df) 
```

```{r}
# Descriptive statistics 
summary(df)
```

```{r}
# Split train : test as 70 : 30
set.seed(6)
train_indices <- createDataPartition(df$Outcome, p = 0.7, list = FALSE)
test_indices <- setdiff(1:nrow(df), train_indices)
# If replace = True, the same data point can appear in both the training and test set

train  <- df[train_indices, ]
test   <- df[test_indices, ]

X_train <- df[train_indices, 1:8]
y_train <- df[train_indices, 9]
X_test <- df[test_indices, 1:8]
y_test <- df[test_indices, 9]

# Examine the size of the training and test sets
cat("Training set's size:", nrow(train), "\n")
cat("Test set's size:", nrow(test), "\n")
```

```{r fig.width=14, fig.height=12}
# # When checking outliers through boxplot / Rosner test / Mahalanobis distance, you only need to check outliers for the numerical feature variables, not the outcome variable

# Check for outliers in numerical feature variables of the training set 
boxplot(train[, 1:8], na.action = NULL)  # na.action = NULL means to ignore NA values in the dataset 
# => From the boxplot, it seems like Pregnancies, Glucose, BloodPressure, SkinThickness, Insulin, BMI, DiabetesPedigreeFunction, Age may have outliers
```

```{r}
# Use Rosner test to re - check for outliers on these variables
rosnerTest(train$Pregnancies, k = 2)  # k is the maximum number of suspected outlier
rosnerTest(train$Glucose, k = 4)  
rosnerTest(train$BloodPressure, k = 11)
rosnerTest(train$SkinThickness, k = 2)
rosnerTest(train$Insulin, k = 11)
rosnerTest(train$BMI, k = 9)
rosnerTest(train$DiabetesPedigreeFunction, k = 7)
rosnerTest(train$Age, k = 2)
# => Pregnancies, SkinThickness and Age each has 1 outlier. Glucose has 3 outliers. BMI has 8 outliers. DiabetesPedigreeFunction has 6 outliers. For BloodPressure and Insulin, the test indicates that each of them has at least 11 outliers. However, the Rosner test doesn't run simulations for k > 10, so we can't be certain about the number of outliers that BloodPressure and Insulin have
```

```{r}
# Find columns with NA values
colSums(is.na(train))[colSums(is.na(train))>0]
# => No missing values 
```

```{r}
## Check for class distribution in training set
count <- as.numeric(table(y_train))
custom_labels <- ifelse(names(table(y_train)) == "1", "Diabetic", "Not Diabetic") 
bp <- barplot(table(y_train), names.arg = custom_labels, main = "Class distribution in Training Set", col = c("aquamarine3", "lightcoral"), ylim = c(0, 400), xlab = "Disease Status", ylab = "Number of Patients")
text(x = bp, y = count, label = count, cex = 0.9, pos = 3, col = "black")
# -> There is only mild class imbalance (around 65:35 ratio between the two classes)
```

```{r}
# Only standardize numerical feature variables, not the outcome variable
X_test_unscaled <- X_test 

# Standardize the training set
X_train <- scale(X_train, center = TRUE, scale = TRUE)

# Apply the same standardization method on the test set 
train_mean <- attr(X_train, "scaled:center")  # Same as colMeans(X_train)
train_sd <- attr(X_train, "scaled:scale")     # Same as apply(X_train, 2, sd)

X_test <- scale(X_test, center = train_mean, scale= train_sd)

# Change format from matrix to dataframe 
X_train <- as.data.frame(X_train)
train <- cbind(X_train, Outcome = y_train)

X_test <- as.data.frame(X_test)
```

### A/ Find training and test set AUC for different classification methods:

#### 1. Logistic regression model:

```{r}
# Fit the logistic regression model
glm_model <- glm(Outcome ~ ., data = train, family = binomial(link = "logit"))
```

```{r}
# Show model's coefficients
summary(glm_model)
```

```{r}
y_glm_train_pred = predict.glm(glm_model, X_train, type = 'response')  # y_pred = P(Y = 1 | X)
y_glm_test_pred = predict.glm(glm_model, X_test, type = 'response')

# Compute the AUC and 95% confidence interval of the AUC for the training set
roc_train <- pROC::roc(response = y_train, predictor = y_glm_train_pred)
auc_train <- pROC::auc(roc_train)
ci_train <- ci.auc(roc_train, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

# Compute the AUC and 95% confidence interval of the AUC for the test set
roc_test <- pROC::roc(y_test, y_glm_test_pred)
auc_test <- pROC::auc(roc_test)
ci_test <- ci.auc(roc_test, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

cat("AUC on training set: ", auc_train, "\n")
cat("95% CI of AUC on training set: [", ci_train[1], " ; ", ci_train[3], "] \n")
cat("AUC on test set: ", auc_test, "\n")
cat("95% CI of AUC on test set: [", ci_test[1], " ; ", ci_test[3], "] \n")
```

```{r}
# Array of false positive rates, true positive rates and thresholds 
fpr_glm_test <- 1 - roc_test$specificities
tpr_glm_test <- roc_test$sensitivities
thresholds_glm_test <- roc_test$thresholds 
# fpr[i] and tpr[i] are the false positive rate and true positive rate at the threshold value thresholds[i]
```

#### **2. Regularized Logistic regression model:**

```{r}
# Fit the logistic regression model with LASSO regularization
glm_lasso_model <- cv.glmnet(x = as.matrix(X_train), y = y_train, alpha = 1, family = "binomial")
# alpha = 1 for LASSO (L1 regularization); family = "binomial" uses the logit link; cv.glmnet uses cross validation to find the optimal lambda

# Show model's coefficients
glm_lasso_coef <- coef(glm_lasso_model, s = "lambda.min")  # Obtain the coefficients
glm_lasso_coef
```

```{r}
y_glm_lasso_train_pred = predict(glm_lasso_model, as.matrix(X_train), s = "lambda.min", type = 'response')  # y_pred = P(Y = 1 | X)
y_glm_lasso_test_pred = predict(glm_lasso_model, as.matrix(X_test), s = "lambda.min", type = 'response')
# -> Predictions use the optimal lambda found by cross validation 

# Compute the AUC and 95% confidence interval of the AUC for the training set
roc_train <- pROC::roc(y_train, as.vector(y_glm_lasso_train_pred))
auc_train <- pROC::auc(roc_train)
ci_train <- ci.auc(roc_train, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

# Compute the AUC and 95% confidence interval of the AUC for the test set
roc_test <- pROC::roc(y_test, as.vector(y_glm_lasso_test_pred))
auc_test <- pROC::auc(roc_test)
ci_test <- ci.auc(roc_test, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

cat("AUC on training set: ", auc_train, "\n")
cat("95% CI of AUC on training set: [", ci_train[1], " ; ", ci_train[3], "] \n")
cat("AUC on test set: ", auc_test, "\n")
cat("95% CI of AUC on test set: [", ci_test[1], " ; ", ci_test[3], "] \n")
```

```{r}
# Array of false positive rates and true positive rates 
fpr_glm_lasso_test <- 1 - roc_test$specificities
tpr_glm_lasso_test <- roc_test$sensitivities
```

#### **3. Generalized Additive Model:**

```{r}
# Fit the GAM model
gam_model <- gam(Outcome ~ s(Pregnancies, k = 15, bs = 'cr') + 
                   s(Glucose, k = 15, bs = 'cr') + 
                   s(BloodPressure, k = 15, bs = 'cr') + 
                   s(SkinThickness, k = 15, bs = 'cr') +
                   s(Insulin   , k = 15, bs = 'cr') + 
                   s(BMI, k = 15, bs = 'cr') + 
                   s(DiabetesPedigreeFunction, k = 15, bs = 'cr') + 
                   s(Age, k = 15, bs = 'cr'),
                 family = binomial(link = "logit"), data = train, method = 'REML', fit = TRUE) 

summary(gam_model)
```

```{r}
# Plot GAM model 
par(mfrow = c(2,4))
plot(gam_model)
```

```{r}
y_gam_train_pred = predict.gam(gam_model, X_train, type = 'response')  # y_pred = P(Y = 1 | X)
y_gam_test_pred = predict.gam(gam_model, X_test, type = 'response')

# Compute the AUC and 95% confidence interval of the AUC for the training set
roc_train <- pROC::roc(y_train, y_gam_train_pred)
auc_train <- pROC::auc(roc_train)
ci_train <- ci.auc(roc_train, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

# Compute the AUC and 95% confidence interval of the AUC for the test set
roc_test <- pROC::roc(y_test, y_gam_test_pred)
auc_test <- pROC::auc(roc_test)
ci_test <- ci.auc(roc_test, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

cat("AUC on training set: ", auc_train, "\n")
cat("95% CI of AUC on training set: [", ci_train[1], " ; ", ci_train[3], "] \n")
cat("AUC on test set: ", auc_test, "\n")
cat("95% CI of AUC on test set: [", ci_test[1], " ; ", ci_test[3], "] \n")
```

```{r}
# Array of false positive rates and true positive rates 
fpr_gam_test <- 1 - roc_test$specificities
tpr_gam_test <- roc_test$sensitivities
```

#### **4. Decision Tree model:**

```{r}
# Fit the initial decision tree model
dtree_model <- rpart(Outcome ~ ., data = train, method = "class", parms = list(split = "gini"), control = rpart.control(cp = 0.006)) # cp is the complexity parameter 
printcp(dtree_model)
```

```{r}
# Plot decision tree
rpart.plot(dtree_model)
```

```{r}
y_dtree_train_pred = predict(dtree_model, X_train, type = 'prob')[,2]  # y_pred = P(Y = 1 | X)
y_dtree_test_pred = predict(dtree_model, X_test, type = 'prob')[,2]

# Compute the AUC and 95% confidence interval of the AUC for the training set
roc_train <- pROC::roc(y_train, y_dtree_train_pred)
auc_train <- pROC::auc(roc_train)
ci_train <- ci.auc(roc_train, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

# Compute the AUC and 95% confidence interval of the AUC for the test set
roc_test <- pROC::roc(y_test, y_dtree_test_pred)
auc_test <- pROC::auc(roc_test)
ci_test <- ci.auc(roc_test, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

cat("AUC on training set: ", auc_train, "\n")
cat("95% CI of AUC on training set: [", ci_train[1], " ; ", ci_train[3], "] \n")
cat("AUC on test set: ", auc_test, "\n")
cat("95% CI of AUC on test set: [", ci_test[1], " ; ", ci_test[3], "] \n")
```

```{r}
# Array of false positive rates and true positive rates
fpr_dtree_test <- 1 - roc_test$specificities
tpr_dtree_test <- roc_test$sensitivities
```

#### **5. Random Forest model:**

```{r}
# Fit the Random Forest model
rf_model <- randomForest(X_train, as.factor(y_train), importance=TRUE, ntree = 500, replace = TRUE)
print(rf_model)
```

```{r}
y_rf_train_pred = predict(rf_model, X_train, type = 'prob')[,2]  # y_pred = P(Y = 1 | X)
y_rf_test_pred = predict(rf_model, X_test, type = 'prob')[,2]

# Compute the AUC and 95% confidence interval of the AUC for the training set
roc_train <- pROC::roc(y_train, y_rf_train_pred)
auc_train <- pROC::auc(roc_train)
ci_train <- ci.auc(roc_train, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

# Compute the AUC and 95% confidence interval of the AUC for the test set
roc_test <- pROC::roc(y_test, y_rf_test_pred)
auc_test <- pROC::auc(roc_test)
ci_test <- ci.auc(roc_test, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

cat("AUC on training set: ", auc_train, "\n")
cat("95% CI of AUC on training set: [", ci_train[1], " ; ", ci_train[3], "] \n")
cat("AUC on test set: ", auc_test, "\n")
cat("95% CI of AUC on test set: [", ci_test[1], " ; ", ci_test[3], "] \n")
```

```{r}
# Array of false positive rates and true positive rates
fpr_rf_test <- 1 - roc_test$specificities
tpr_rf_test <- roc_test$sensitivities
```

#### **6.** Extreme Gradient Boosting **model:**

```{r}
# Fit the XGBoost model
xgb_model <- xgboost(data = as.matrix(X_train), label = as.numeric(as.character(y_train)), nrounds = 100, eta = 0.06, objective = "binary:logistic", verbose = 0) # gamma is the regularization parameter for tree based models  

# Plot the third individual decision tree in the model
xgb.plot.tree(model = xgb_model, trees = 3, render = TRUE)
```

```{r}
y_xgb_train_pred = predict(xgb_model, as.matrix(X_train), type = 'prob')  # y_pred = P(Y = 1 | X)
y_xgb_test_pred = predict(xgb_model, as.matrix(X_test), type = 'prob')

# Compute the AUC and 95% confidence interval of the AUC for the training set
roc_train <- pROC::roc(y_train, y_xgb_train_pred)
auc_train <- pROC::auc(roc_train)
ci_train <- ci.auc(roc_train, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

# Compute the AUC and 95% confidence interval of the AUC for the test set
roc_test <- pROC::roc(y_test, y_xgb_test_pred)
auc_test <- pROC::auc(roc_test)
ci_test <- ci.auc(roc_test, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

cat("AUC on training set: ", auc_train, "\n")
cat("95% CI of AUC on training set: [", ci_train[1], " ; ", ci_train[3], "] \n")
cat("AUC on test set: ", auc_test, "\n")
cat("95% CI of AUC on test set: [", ci_test[1], " ; ", ci_test[3], "] \n")
```

```{r}
# Array of false positive rates and true positive rates 
fpr_xgb_test <- 1 - roc_test$specificities
tpr_xgb_test <- roc_test$sensitivities
```

#### **7. Support Vector Machine model:**

```{r}
# Fit the SVM model
svm_model <- svm(Outcome ~ ., data = train, type = "C-classification", kernel = "linear", probability = TRUE, cost = 0.6)
```

```{r}
y_svm_train_pred = attr(predict(svm_model, X_train, probability = TRUE), "probabilities")[,1]  # y_pred = P(Y = 1 | X)
y_svm_test_pred = attr(predict(svm_model, X_test, probability = TRUE), "probabilities")[,1]

# Compute the AUC and 95% confidence interval of the AUC for the training set
roc_train <- pROC::roc(y_train, y_svm_train_pred)
auc_train <- pROC::auc(roc_train)
ci_train <- ci.auc(roc_train, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

# Compute the AUC and 95% confidence interval of the AUC for the test set
roc_test <- pROC::roc(y_test, y_svm_test_pred)
auc_test <- pROC::auc(roc_test)
ci_test <- ci.auc(roc_test, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

cat("AUC on training set: ", auc_train, "\n")
cat("95% CI of AUC on training set: [", ci_train[1], " ; ", ci_train[3], "] \n")
cat("AUC on test set: ", auc_test, "\n")
cat("95% CI of AUC on test set: [", ci_test[1], " ; ", ci_test[3], "] \n")
```

```{r}
# Array of false positive rates and true positive rates
fpr_svm_test <- 1 - roc_test$specificities
tpr_svm_test <- roc_test$sensitivities
```

#### **8. K - Nearest Neighbor model:**

```{r}
# Fit the KNN classifier model
knn_model <- knn3(Outcome ~ ., data = train, k = 45)
```

```{r}
y_knn_train_pred = predict(knn_model, X_train, type = 'prob')[,2]  # y_pred = P(Y = 1 | X)  # y_pred = P(Y = 1 | X)
y_knn_test_pred = predict(knn_model, X_test, type = 'prob')[,2]

# Compute the AUC and 95% confidence interval of the AUC for the training set
roc_train <- pROC::roc(y_train, y_knn_train_pred)
auc_train <- pROC::auc(roc_train)
ci_train <- ci.auc(roc_train, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

# Compute the AUC and 95% confidence interval of the AUC for the test set
roc_test <- pROC::roc(y_test, y_knn_test_pred)
auc_test <- pROC::auc(roc_test)
ci_test <- ci.auc(roc_test, method = "bootstrap", conf.level = 0.95, boot.n = 2000)

cat("AUC on training set: ", auc_train, "\n")
cat("95% CI of AUC on training set: [", ci_train[1], " ; ", ci_train[3], "] \n")
cat("AUC on test set: ", auc_test, "\n")
cat("95% CI of AUC on test set: [", ci_test[1], " ; ", ci_test[3], "] \n")
```

```{r}
# Array of false positive rates and true positive rates
fpr_knn_test <- 1 - roc_test$specificities
tpr_knn_test <- roc_test$sensitivities
```

### **B/ Choosing decision threshold by ROC curve and decision curve:**

#### 1. Plot ROC curve of each method:

```{r fig.width=6, fig.height=5.5}
plot(0, type="n", main="ROC curve for each method", xlim=c(0, 1), ylim=c(0, 1), xlab="False positive rate",ylab="True positive rate")
lines(fpr_glm_test, tpr_glm_test, col="blue", type="l", lty=1, lwd=3)
lines(fpr_glm_lasso_test, tpr_glm_lasso_test, col="red", type="l", lty=1, lwd=3)
lines(fpr_gam_test, tpr_gam_test, col="green", type="l", lty=1, lwd=3)
lines(fpr_dtree_test, tpr_dtree_test, col="purple", type="l", lty=1, lwd=3)
lines(fpr_rf_test, tpr_rf_test, col="orange", type="l", lty=1, lwd=3)
lines(fpr_xgb_test, tpr_xgb_test, col="brown", type="l", lty=1, lwd=3)
lines(fpr_svm_test, tpr_svm_test, col="darkcyan", type="l", lty=1, lwd=3)
lines(fpr_knn_test, tpr_knn_test, col="burlywood3", type="l", lty=1, lwd=3)
legend("bottomright", legend=c("Logistic Reg", "R. Logistic Reg", "GAM", "Decision Tree", "RF", "XGBoost", "SVM", "KNN"), col=c("blue", "red", "green", "purple", "orange", "brown", "darkcyan", "burlywood3"), lty=1, lwd=3)
```

#### 2. Plot decision curve of each method:

```{r}
df_ <- data.frame(y_glm_test_pred, y_glm_lasso_test_pred, y_gam_test_pred, y_dtree_test_pred, y_rf_test_pred, y_xgb_test_pred, y_svm_test_pred, y_knn_test_pred, y_test)

colnames(df_)[2] <- 'y_glm_lasso_test_pred'

# Calculate disease prevalence in the population:
p = sum(test$Outcome == 1)/nrow(test)

dcurves::dca(y_test ~ ., data = df_, label = list(y_glm_test_pred = "Logistic Reg", y_glm_lasso_test_pred = "R. Logistic Reg", y_gam_test_pred = "GAM", y_dtree_test_pred = "Decision Tree", y_rf_test_pred = "RF", y_xgb_test_pred = "XGBoost", y_svm_test_pred = "SVM", y_knn_test_pred = "KNN"), prevalence = p) %>% 
  plot(smooth = FALSE) + geom_line(linewidth = 1) + labs(title = "Decision curve for each method") +
  theme(plot.title = element_text(face="bold", hjust = 0.5))
```

#### **3. Criterias to determine the decision threshold for the Logistic regression method:**

-   **Youden index criterion:**

```{r}
# Optimal threshold obtained by maximizing youden index
j_threshold <- function(y_true, y_pred) {
  roc_obj <- pROC::roc(y_true, y_pred)
  fpr <- 1 - roc_obj$specificities
  tpr <- roc_obj$sensitivities
  thresholds <- roc_obj$thresholds
  j <- tpr - fpr 
  j <- sort(j, decreasing = TRUE, index.return = TRUE)
  j_ordered <- data.frame(j = j[1], thresholds = thresholds[rank(j$ix)])
  for (k in 1:nrow(j_ordered)){
    if (is.finite(j_ordered$thresholds[k])) {
      return(j_ordered$thresholds[k])
      break
      }
  }
}

cj_glm <- j_threshold(y_test, y_glm_test_pred)
cj_glm
```

-   **Closest to (0,1) criterion:**

```{r}
# Optimal threshold obtained by maximizing sensitivity and specificity to the extent that both are closest to 100% 
d_threshold <- function(y_true, y_pred) {
  roc_obj <- pROC::roc(y_true, y_pred)
  fpr <- 1 - roc_obj$specificities
  tpr <- roc_obj$sensitivities
  thresholds <- roc_obj$thresholds
  d <- sapply(1:length(fpr), function(i) sqrt((1 - tpr[i])^2 + (fpr[i])^2))
  d <- sort(d, decreasing = FALSE, index.return = TRUE)
  d_ordered <- data.frame(d = d[1], thresholds = thresholds[rank(d$ix)])
  for (k in 1:nrow(d_ordered)){
    if (is.finite(d_ordered$thresholds[k])) {
      return(d_ordered$thresholds[k])
      break
      }
  }
}

cd_glm <- d_threshold(y_test, y_glm_test_pred)
cd_glm
```

-   **Symetric point criterion:**

```{r}
# Optimal threshold obtained by finding the point where sensitivity and specificity are equivalent 
s_threshold <- function(y_true, y_pred) {
  roc_obj <- pROC::roc(y_true, y_pred)
  fpr <- 1 - roc_obj$specificities
  tpr <- roc_obj$sensitivities
  thresholds <- roc_obj$thresholds
  s <- abs(fpr + tpr - 1)
  s <- sort(s, decreasing = FALSE, index.return = TRUE)
  s_ordered <- data.frame(s = s[1], thresholds = thresholds[rank(s$ix)])
  for (k in 1:nrow(s_ordered)){
    if (is.finite(s_ordered$thresholds[k])) {
      return(s_ordered$thresholds[k])
      break
      }
  }
}

cs_glm <- s_threshold(y_test, y_glm_test_pred)
cs_glm
```

#### 4. Plot ROC curve, decision curve and optimal thresholds for the **Logistic regression** method:

```{r}
# Plot ROC curve:
roc_obj <- pROC::roc(y_test, y_glm_test_pred)
thresholds_glm <- roc_obj$thresholds 
roc_data <- data.frame(fpr = rev(1 - roc_obj$specificities), tpr = rev(roc_obj$sensitivities))

for (i in 1:length(roc_obj$thresholds)){
  if (roc_obj$thresholds[i] == cj_glm){
    specificity_j <- roc_obj$specificities[i]
    sensitivity_j <- roc_obj$sensitivities[i]
  }
  if (roc_obj$thresholds[i] == cd_glm){
    specificity_d <- roc_obj$specificities[i]
    sensitivity_d <- roc_obj$sensitivities[i]
  }
  if (roc_obj$thresholds[i] == cs_glm){
    specificity_s <- roc_obj$specificities[i]
    sensitivity_s <- roc_obj$sensitivities[i]
  }
}

diagonal_data <- data.frame(x = seq(0, 1, length.out = 100), y = 1 - seq(0, 1, length.out = 100))
thresholds_data <- data.frame(x1 = 0, x2 = 1 - specificity_d, y1 = 1, y2 = sensitivity_d)
sp_se_j <- data.frame(sp = specificity_j, se = sensitivity_j)
sp_se_d <- data.frame(sp = specificity_d, se = sensitivity_d)
sp_se_s <- data.frame(sp = specificity_s, se = sensitivity_s)

ggplot(data = roc_data, aes(x = fpr, y = tpr)) + 
  geom_line(linetype = "solid", linewidth = 1, col = "blue") +
  geom_line(data = diagonal_data, aes(x = x, y = y), linetype = "dashed", linewidth = 1, col = "orange") +
  geom_segment(data = thresholds_data, aes(x = x1, y = y1, xend = x2, yend = y2), linetype = "dotdash", linewidth = 1, col = "pink") +
  geom_point(data = sp_se_j, aes(x = 1 - sp, y = se, col = "c_j"), size = 4, show.legend = TRUE) +
  geom_point(data = sp_se_d, aes(x = 1 - sp, y = se, col = "c_d"), size = 4, show.legend = TRUE) +
  geom_point(data = sp_se_s, aes(x = 1 - sp, y = se, col = "c_s"), size = 4, show.legend = TRUE) +
  labs(title = "ROC curve", x = "False positive rate", y = "True positive rate") + 
  scale_color_manual(name = "Threshold", labels = unname(TeX(c("At $\\hat{c}_{J}$", "At $\\hat{c}_{D}$", "At $\\hat{c}_{S}$"))), values=c("red", "purple", "green")) +
  theme_bw() +
  theme(plot.title = element_text(face="bold", hjust = 0.5)) 
```

```{r}
# Plot decision curve

# Net benefit function
nb <- function(y_true, y_pred, c, prev){
  roc_obj <- pROC::roc(y_true, y_pred)
  for (i in 1:length(roc_obj$thresholds)){
    if (roc_obj$thresholds[i] == c){
      se_c <- roc_obj$sensitivities[i]
      sp_c <- roc_obj$specificities[i]
      break
    }
  }
  phi_c <- se_c * prev - (1 - sp_c) * (1 - prev) * c/(1 - c) 
  return(phi_c)
}
  
df_glm <- data.frame(y_glm_test_pred, y_test)
nb_j <- data.frame(cj = cj_glm, nbj = nb(y_test, y_glm_test_pred, cj_glm, p))
nb_d <- data.frame(cd = cd_glm, nbd = nb(y_test, y_glm_test_pred, cd_glm, p))
nb_s <- data.frame(cs = cs_glm, nbs = nb(y_test, y_glm_test_pred, cs_glm, p))

# For the Decision curve plot, use the thresholds values from ROC curve plot to make sure that the optimal thresholds show on the decision curve correctly
dcurves::dca(y_test ~ y_glm_test_pred, data = df_glm, label = list(y_glm_test_pred = "GLM"), thresholds = thresholds_glm, prevalence = p) %>% 
  plot(smooth = FALSE) + geom_line(linewidth = 1) + labs(title = "Decision curve") +
  geom_point(data = nb_j, aes(x = cj, y = nbj, col = "c_j"), size = 4, show.legend = TRUE) +
  geom_point(data = nb_d, aes(x = cd, y = nbd, col = "c_d"), size = 4, show.legend = TRUE) +
  geom_point(data = nb_s, aes(x = cs, y = nbs, col = "c_s"), size = 4, show.legend = TRUE) + 
  scale_color_manual(labels = unname(TeX(c("Treat All", "Treat None", "GLM", "At $\\hat{c}_{J}$", "At $\\hat{c}_{D}$", "At $\\hat{c}_{S}$"))), values=c("orange", "pink", "blue", "red", "purple", "green")) +
  guides(colour = guide_legend(override.aes = list(linetype = c("solid", "solid", "solid", rep("blank", 3)), shape = c(NA, NA, NA, rep(16, 3))))) + 
  theme(plot.title = element_text(face="bold", hjust = 0.5))
```

```{r}
# Finding the optimal threshold
nb_j <- nb(y_test, y_glm_test_pred, cj_glm, p)
nb_d <- nb(y_test, y_glm_test_pred, cd_glm, p)
nb_s <- nb(y_test, y_glm_test_pred, cs_glm, p)
nb_values <- c(cj = nb_j, cd = nb_d, cs = nb_s)

best_threshold <- names(which.max(nb_values))   # Name of the threshold with highest net benefit
threshold_map <- list(cj = cj_glm, cd = cd_glm, cs = cs_glm)
c_res <- threshold_map[[best_threshold]]   # Get the optimal threshold's value 
cat("Optimal threshold:", best_threshold, "=", c_res)
```

```{r}
## Example: Predicting the class of a new data point with GLM model
new_obs_ <- X_test_unscaled[32:37,]  # Some observations from the test set
new_obs <- scale(new_obs_, center = train_mean, scale= train_sd)
new_obs <- as.data.frame(new_obs)
# In the procedure: You need to preprocess (Convert categorical variables into factors, Fill in NA values, Standardize, ... in the same method and with the same values calculated from the preprocessing step of the training set) the new data if needed 
actual_classes <- y_test[32:37]
class_preds <- c()

y_glm_pred = predict.glm(glm_model, new_obs, type = 'response')
y_glm_pred = round(y_glm_pred, digits = 7)
class_preds <- ifelse(y_glm_pred >= c_res, "Diabetic", "Not Diabetic")
actual_classes <- ifelse(actual_classes == 1, "Diabetic", "Not Diabetic")

# Show the observations
obs_names <- paste("Observation", 1:6)
obs_names <- data.frame(Observation = unlist(obs_names))
print(cbind(obs_names, new_obs_))

# Final result
res <- data.frame(obs_names, "Predicted probability" = y_glm_pred, "Predicted class" = class_preds, "Actual class" = actual_classes)
res
```
